---
output: html_document
editor_options:
  chunk_output_type: console
css: highlight.css
---

```{r setup, include=FALSE}
knitr::knit_hooks$set(source = function(x, options) {
  if (!is.null(options$hilang)) {
      code_open <- "\n\n<div class=\"sourceCode\">\n<pre class=\"sourceCode\">\n<code class=\"sourceCode\">"
      code_close <- "\n</code>\n</pre>\n</div>\n"
      code_body <- highr::hi_andre(x, language = options$hilang, format = "html")
    stringr::str_c(
      code_open,
      knitr:::indent_block(paste(code_body, collapse = '\n'), ""),
      code_close
    )
  } else {
    stringr::str_c("\n\n```", tolower(options$engine), "\n",
                   paste(x, collapse = '\n'), "\n```\n\n")

  }
})
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')
```


```{r, include=FALSE}
# source('R/draw_workflow.R')
# prod_het <- tibbler(c('inds_out', 'win_out'))
```

# (git 11) Visualization

After all `nextflow` pipelines are successfully run to completion, each Figure (and Suppl. Figure) of the manuscript can be re-created with its respective `R` script located under `R/fig`.
These are executable `R` scripts that can be launched from the base directory;

```sh
Rscript --vanilla R/fig/plot_Fxyz.R input1 input2 ...
```

For convenience, there also exists a `bash` script that can be used to re-create all Figures in one go (git 11):

```sh
cd $BASE_DIR
bash sh/create_figures.sh
```

After running `create_figures.sh`, Figures 1 - 5 and Suppl. Figures 1 - 9 should be created withing the folder `figures/`.

In the remaining documentation, the individual Visualization scripts are going to discussed in detail.

---
