---
output: html_document
editor_options:
  chunk_output_type: console
css: highlight.css
---

```{r setup, include=FALSE}
knitr::knit_hooks$set(source = function(x, options) {
  if (!is.null(options$hilang)) {
      code_open <- "\n\n<div class=\"sourceCode\">\n<pre class=\"sourceCode\">\n<code class=\"sourceCode\">"
      code_close <- "\n</code>\n</pre>\n</div>\n"
      code_body <- highr::hi_andre(x, language = options$hilang, format = "html")
    stringr::str_c(
      code_open,
      knitr:::indent_block(paste(code_body, collapse = '\n'), ""),
      code_close
    )
  } else {
    stringr::str_c("\n\n```", tolower(options$engine), "\n",
                   paste(x, collapse = '\n'), "\n```\n\n")

  }
})
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')
```


```{r, include=FALSE}
# source('R/draw_workflow.R')
# prod_het <- tibbler(c('inds_out', 'win_out'))
```

# (git 15) Data Visualization

After all `nextflow` pipelines are successfully run to completion, each Figure (and Suppl. Figure) of the manuscript can be re-created with its respective `R` script located under `R/fig`.
These are executable `R` scripts that can be launched from the base directory;

```sh
Rscript --vanilla R/fig/plot_Fxyz.R input1 input2 ...
```

For convenience, there also exists a `bash` script that can be used to re-create all Figures in one go (git 15):

```sh
cd $BASE_DIR
bash sh/create_figures.sh
```

After running `create_figures.sh`, Figures 1 - 6 and Suppl. Figures 1 - 16 should be created withing the folder `figures/`.

In the remaining documentation, the individual Visualization scripts are going to discussed in detail.

Below is the bash code that is executed when running `create_figures.sh`:

```sh
#/usr/bin/bash
# git 15

# Main Figures

Rscript --vanilla R/fig/plot_F1.R \
  2_analysis/fst/50k/ \
  2_analysis/summaries/fst_globals.txt \
  2_analysis/summaries/fst_permutation_summary.tsv

Rscript --vanilla R/fig/plot_F2.R \
  2_analysis/msmc/output/ \
  2_analysis/cross_coalescence/output/ \
  2_analysis/msmc/setup/msmc_grouping.txt \
  2_analysis/msmc/setup/msmc_cc_grouping.txt \
  2_analysis/summaries/fst_globals.txt

Rscript --vanilla R/fig/plot_F3.R \
  2_analysis/fst/50k/ \
  2_analysis/summaries/fst_globals.txt

Rscript --vanilla R/fig/plot_F4.R 2_analysis/dxy/50k/ \
  2_analysis/fst/50k/multi_fst.50k.tsv.gz 2_analysis/GxP/50000/ \
  2_analysis/summaries/fst_outliers_998.tsv \
  https://raw.githubusercontent.com/simonhmartin/twisst/master/plot_twisst.R \
  2_analysis/twisst/weights/ ressources/plugin/trees/ \
  2_analysis/fasteprr/step4/fasteprr.all.rho.txt.gz \
  2_analysis/summaries/fst_globals.txt

Rscript --vanilla R/fig/plot_F5.R \
  2_analysis/twisst/weights/ ressources/plugin/trees/ \
  https://raw.githubusercontent.com/simonhmartin/twisst/master/plot_twisst.R \
  2_analysis/summaries/fst_outliers_998.tsv 2_analysis/dxy/50k/ \
  2_analysis/fst/50k/ 2_analysis/summaries/fst_globals.txt \
  2_analysis/GxP/50000/ 200 5 2_analysis/revPoMo/outlier_regions/

Rscript --vanilla R/fig/plot_F6.R \
  2_analysis/summaries/fst_outliers_998.tsv \
  2_analysis/geva/ 2_analysis/GxP/bySNP/

# Suppl. Figures

Rscript --vanilla R/fig/plot_SF1.R \
  2_analysis/pca/

Rscript --vanilla R/fig/plot_SF2.R \
  2_analysis/dxy/50k/ \
  2_analysis/fst/50k/ \
  2_analysis/summaries/fst_globals.txt

Rscript --vanilla R/fig/plot_SF3.R \
  2_analysis/dxy/50k/

Rscript --vanilla R/fig/plot_SF4.R \
  2_analysis/newhyb/nh_input/NH.Results/

Rscript --vanilla R/fig/plot_SF5.R 2_analysis/fst/50k/ \
  2_analysis/summaries/fst_outliers_998.tsv \
  2_analysis/summaries/fst_globals.txt

Rscript --vanilla R/fig/plot_SF6.R \
  2_analysis/summaries/fst_globals.txt \
  2_analysis/fst/50k/ \
  2_analysis/fasteprr/step4/fasteprr.all.rho.txt.gz

Rscript --vanilla R/fig/plot_SF7.R \
  2_analysis/dxy/50k/

Rscript --vanilla R/fig/plot_SF8.R \
  2_analysis/pi/50k/ \
  2_analysis/fasteprr/step4/fasteprr.all.rho.txt.gz

Rscript --vanilla R/fig/plot_SF9.R \
  2_analysis/raxml/hyp155_n_0.33_mac4_5kb.raxml.support.bs-tbe \
  2_analysis/raxml/RAxML_bipartitions.hypS-h_n_0.33_mac6_10kb

Rscript --vanilla R/fig/plot_SF10.R \
  2_analysis/pi/50k/

Rscript --vanilla R/fig/plot_SF11.R \
  2_analysis/raxml/lg04.1_155N.raxml.support \
  2_analysis/raxml/lg12.3_155N.raxml.support \
  2_analysis/raxml/lg12.4_155N.raxml.support

Rscript --vanilla R/fig/plot_SF12.R \
  2_analysis/raxml/lg04.1_hySN.raxml.support \
  2_analysis/raxml/lg12.3_hySN.raxml.support \
  2_analysis/raxml/lg12.4_hySN.raxml.support

Rscript --vanilla R/fig/plot_SF13.R \
  2_analysis/admixture/ \
  metadata/phenotypes.sc

Rscript --vanilla R/fig/plot_SF14.R \
2_analysis/GxP/50000/

Rscript --vanilla R/fig/plot_SF15.R \
  2_analysis/fst_signif/random/

Rscript --vanilla R/fig/plot_SF16.R \
  2_analysis/raxml/hyS_n_0.33_mac4_5kb.raxml.support
# ==================
```

---
