---
output: html_document
editor_options:
  chunk_output_type: console
css: highlight.css
---

```{r setup, include=FALSE}
knitr::knit_hooks$set(source = function(x, options) {
  if (!is.null(options$hilang)) {
      code_open <- "\n\n<div class=\"sourceCode\">\n<pre class=\"sourceCode\">\n<code class=\"sourceCode\">"
      code_close <- "\n</code>\n</pre>\n</div>\n"
      code_body <- highr::hi_andre(x, language = options$hilang, format = "html")
    stringr::str_c(
      code_open,
      knitr:::indent_block(paste(code_body, collapse = '\n'), ""),
      code_close
    )
  } else {
    stringr::str_c("\n\n```", tolower(options$engine), "\n",
                   paste(x, collapse = '\n'), "\n```\n\n")

  }
})
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')
```


```{r, include=FALSE}
# source('R/draw_workflow.R')
# prod_het <- tibbler(c('inds_out', 'win_out'))
```

# (git 20) Data Visualization

After all `nextflow` pipelines are successfully run to completion, each Figure (and Suppl. Figure) of the manuscript can be re-created with its respective `R` script located under `R/fig`.
These are executable `R` scripts that can be launched from the base directory;

```sh
Rscript --vanilla R/fig/plot_Fxyz.R input1 input2 ...
```

For convenience, there also exists a `bash` script that can be used to re-create all Figures in one go (git 15):

```sh
cd $BASE_DIR
bash sh/create_figures.sh
```

After running `create_figures.sh`, Figures 1 - 6 and Suppl. Figures 1 - 21 should be created withing the folder `figures/`.

In the remaining documentation, the individual Visualization scripts are going to discussed in detail.

Below is the bash code that is executed when running `create_figures.sh`:

```sh
#/usr/bin/bash
# git 20

# preparation for Fig. 1
Rscript --vanilla R/fst_permutation.R \
    2_analysis/fst_signif/random/

# Main Figures
# -------------------------------------------------
# git 20.1
# needs to be run interactively because fonts are not accessible for Rscript
# Rscript --vanilla R/fig/plot_F1.R \
#  2_analysis/fst/50k/ \
#  2_analysis/summaries/fst_globals.txt \
#  2_analysis/summaries/fst_permutation_summary.tsv \
#  2_analysis/fotl/concat_R24ed.treefile

# git 20.2
Rscript --vanilla R/fig/plot_F2.R \
    2_analysis/fst/50k/multi_fst.50k.tsv.gz \
    2_analysis/GxP/50000/ \
    2_analysis/summaries/fst_outliers_998.tsv \
    https://raw.githubusercontent.com/simonhmartin/twisst/master/plot_twisst.R \
    2_analysis/twisst/weights/ \
    ressources/plugin/trees/ \
    2_analysis/summaries/fst_globals.txt

# git 20.3
Rscript --vanilla R/fig/plot_F3.R \
    2_analysis/msmc/output/ \
    2_analysis/cross_coalescence/output/ \
    2_analysis/msmc/setup/msmc_grouping.txt \
    2_analysis/msmc/setup/msmc_cc_grouping.txt \
    2_analysis/summaries/fst_globals.txt

# git 20.4
Rscript --vanilla R/fig/plot_F4.R \
    2_analysis/astral/astral_5000x_5kb_v1_noS.tre \
    2_analysis/ibd/cM_converted/no_outgr_bed95_8.conv_filterd.tsv

# git 20.5
Rscript --vanilla R/fig/plot_F5.R \
    2_analysis/twisst/weights/ \
    ressources/plugin/trees/ \
    https://raw.githubusercontent.com/simonhmartin/twisst/master/plot_twisst.R \
    2_analysis/summaries/fst_outliers_998.tsv \
    2_analysis/dxy/50k/ \
    2_analysis/fst/50k/ \
    2_analysis/summaries/fst_globals.txt \
    2_analysis/GxP/50000/ \
    200 \
    5 \
    2_analysis/revPoMo/outlier_regions/

# git 20.6
Rscript --vanilla R/fig/plot_F6.R \
    2_analysis/summaries/fst_outliers_998.tsv \
    2_analysis/geva/ \
    2_analysis/GxP/bySNP/

# Suppl. Figures
# -------------------------------------------------
# git 20.7
Rscript --vanilla R/fig/plot_SF1.R \
    ressources/Rabosky_etal_2018/dataFiles/ratemat_enhanced.csv

# git 20.8
# needs to be run interactively because fonts are not accessible for Rscript
# Rscript --vanilla R/fig/plot_SF2.R \
#    ressources/Rabosky_etal_2018/

# git 20.9
Rscript --vanilla R/fig/plot_SF3.R \
    2_analysis/pca/

# git 20.10
Rscript --vanilla R/fig/plot_SF4.R \
    2_analysis/fst/50k/ \
    2_analysis/summaries/fst_outliers_998.tsv \
    2_analysis/summaries/fst_globals.txt

# git 20.11
Rscript --vanilla R/fig/plot_SF5.R \
    2_analysis/fst/50k/ \
    2_analysis/summaries/fst_globals.txt

# git 20.12
Rscript --vanilla R/fig/plot_SF6.R \
    2_analysis/summaries/fst_globals.txt \
    2_analysis/fst/50k/ \
    2_analysis/fasteprr/step4/fasteprr.all.rho.txt.gz

# git 20.13
Rscript --vanilla R/fig/plot_SF7.R \
    2_analysis/dxy/50k/

# git 20.14
Rscript --vanilla R/fig/plot_SF8.R \
    2_analysis/dxy/50k/ \
    2_analysis/fst/50k/multi_fst.50k.tsv.gz \
    2_analysis/GxP/50000/ \
    2_analysis/summaries/fst_outliers_998.tsv \
    https://raw.githubusercontent.com/simonhmartin/twisst/master/plot_twisst.R \
    2_analysis/twisst/weights/ \
    ressources/plugin/trees/ \
    2_analysis/fasteprr/step4/fasteprr.all.rho.txt.gz \
    2_analysis/summaries/fst_globals.txt

# git 20.15
Rscript --vanilla R/fig/plot_SF9.R \
    2_analysis/pi/50k/ \
    2_analysis/fasteprr/step4/fasteprr.all.rho.txt.gz

# git 20.16
Rscript --vanilla R/fig/plot_SF10.R \
    2_analysis/dxy/50k/

# git 20.17
Rscript --vanilla R/fig/plot_SF11.R \
    2_analysis/newhyb/nh_input/NH.Results/

# git 20.18
Rscript --vanilla R/fig/plot_SF12.R \
    ressources/species_order_alpha.txt \
    2_analysis/dstats/hyp_ld05_dtrios_BBAA.txt \
    2_analysis/dstats/BBAA_ld05.csv \
    2_analysis/dstats/BBAA_sign_ld05.csv

# git 20.19
Rscript --vanilla R/fig/plot_SF13.R \
    2_analysis/summaries/fst_outliers_998.tsv

# git 20.20
Rscript --vanilla R/fig/plot_SF14.R \
    2_analysis/raxml/lg04.1_155N.raxml.support \
    2_analysis/raxml/lg12.3_155N.raxml.support \
    2_analysis/raxml/lg12.4_155N.raxml.support

# git 20.21
Rscript --vanilla R/fig/plot_SF15.R \
    2_analysis/raxml/lg04.1_hySN.raxml.support \
    2_analysis/raxml/lg12.3_hySN.raxml.support \
    2_analysis/raxml/lg12.4_hySN.raxml.support

# git 20.22
Rscript --vanilla R/fig/plot_SF16.R \
    2_analysis/admixture/ \
    metadata/phenotypes.sc

# git 20.23
Rscript --vanilla R/fig/plot_SF17.R \
    2_analysis/astral/astral_5000x_5kb_v1_all.tre

# git 20.24
Rscript --vanilla R/fig/plot_SF18.R \
    2_analysis/summaries/fst_outliers_998.tsv

# git 20.25
Rscript --vanilla R/fig/plot_SF19.R \
    2_analysis/summaries/fst_outliers_998.tsv \
    2_analysis/geva/ \
    2_analysis/GxP/bySNP/

# git 20.26
Rscript --vanilla R/fig/plot_SF20.R \
    2_analysis/GxP/50000/

# git 20.27
Rscript --vanilla R/fig/plot_SF21.R \
    2_analysis/pi/50k/

# -------------------------------------------------
rm Rplots.pdf
```

---
