---
output: html_document
editor_options:
  chunk_output_type: console
css: highlight.css
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(source = function(x, options) {
  if (!is.null(options$hilang)) {
      code_open <- "\n\n<div class=\"sourceCode\">\n<pre class=\"sourceCode\">\n<code class=\"sourceCode\">"
      code_close <- "\n</code>\n</pre>\n</div>\n"
      code_body <- highr::hi_andre(x, language = options$hilang, format = "html")
    stringr::str_c(
      code_open,
      knitr:::indent_block(paste(code_body, collapse = '\n'), ""),
      code_close
    )
  } else {
    stringr::str_c("\n\n```", tolower(options$engine), "\n",
                   paste(x, collapse = '\n'), "\n```\n\n")

  }
})
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')
```


```{r, include = FALSE}
source('R/draw_workflow.R')
prod_basic <- tibbler(c('gxp_lm_smoothing_output, gxp_lmm_smoothing_output', 'gemma_results',
                        'phenotype_file', 'fst_glob', "fst_10k_output","fst_50k", 'fst_logs',
                        'multi_fst_output', 'fst_outlier_output','vcf_multi_fst'))
```

# (git 12) Analysis X (<i>F<sub>ST</sub></i> Permutation Test)

This pipeline can be executed as follows:

```sh
cd $BASE_DIR/nf/12_analysis_fst_signif
source ../sh/nextflow_alias.sh
nf_run_fstsig
```

## Summary

The <span style="color:red;">...</span> are computed within the [**nextflow**](https://www.nextflow.io/) script `analysis_fst_sign.nf` (located under `$BASE_DIR/nf/12_analysis_fst_signif/`).
It takes the <span style="color:red;">...</span> and computes <span style="color:red;">...</span>.
Below is an overview of the steps involved in the analysis.
(The <span style="color:#4DAF4A">green dot</span> indicates the genotype input, <span style="color:#E41A1C">red arrows</span> depict output that is exported for further use.)

<div style="max-width:800px; margin:auto;">
```{r, eval = FALSE, echo = FALSE, warning = FALSE, message = FALSE}
girafe( ggobj = dot_plot(file = 'docs/analysis_basic.dot',
                         git = 3, point_types = prod_basic),
        width_svg = 16, height_svg = 16)

```
</div>

## Details of `analysis_fst_sign.nf`

### Setup

The nextflow script starts by <span style="color:red;">...</span>

:::kclass
```{r , eval = FALSE, hilang = 'nf'}
#!/usr/bin/env nextflow
// git 12.1
// open genotype data
Channel
	.fromFilePairs("../../1_genotyping/4_phased/phased_mac2.vcf.{gz,gz.tbi}")
	.into{ vcf_locations; vcf_adapt }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.2
// prepare location channel
Channel
	.from( "bel", "hon", "pan")
	.set{ locations_ch }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.3
// prepare subset modes (whole genome vs non-diverged regions)
Channel
	.from( "whg", "subset_non_diverged")
	.into{ subset_type_ch; subset_type_ch2 }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.4
// load table with differentiation outlier regions
Channel
	.fromPath( "../../2_analysis/summaries/fst_outliers_998.tsv" )
	.into{ outlier_tab; outlier_tab2 }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.5
// attach genotypes to location
locations_ch
	.combine( vcf_locations )
	.combine( outlier_tab )
	.set{ vcf_location_combo }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.6
// subset vcf by location
process subset_vcf_by_location {
	label "L_20g2h_subset_vcf"

	input:
	set val( loc ), vcfId, file( vcf ), file( outlier_tab ) from vcf_location_combo

	output:
	set val( loc ), file( "${loc}.vcf.gz" ), file( "${loc}.vcf.gz.tbi" ), file( "${loc}.pop" ), file( outlier_tab ) into ( vcf_loc_pair1, vcf_loc_pair2, vcf_loc_pair3 )

	script:
	"""
	vcfsamplenames ${vcf[0]} | \
		grep ${loc} | \
		grep -v tor | \
		grep -v tab > ${loc}.pop

	vcftools --gzvcf ${vcf[0]} \
		--keep ${loc}.pop \
		--mac 3 \
		--recode \
		--stdout | bgzip > ${loc}.vcf.gz
	
	tabix ${loc}.vcf.gz
	"""
}
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.7
// define location specific sepcies set
Channel.from( [[1, "ind"], [2, "may"], [3, "nig"], [4, "pue"], [5, "uni"]] ).into{ bel_spec1_ch; bel_spec2_ch }
Channel.from( [[1, "abe"], [2, "gum"], [3, "nig"], [4, "pue"], [5, "ran"], [6, "uni"]] ).into{ hon_spec1_ch; hon_spec2_ch }
Channel.from( [[1, "nig"], [2, "pue"], [3, "uni"]] ).into{ pan_spec1_ch; pan_spec2_ch }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.8
// prepare pairwise fsts
// ------------------------------
/* (create all possible species pairs depending on location
   and combine with genotype subset (for the respective location))*/
// ------------------------------
/* channel content after joinig:
  set [0:val(loc), 1:file(vcf), 2:file( vcfidx ), 3:file(pop), 4:file( outlier_tab ), 5:val(spec1), 6:val(spec2)]*/
// ------------------------------
bel_pairs_ch = Channel.from( "bel" )
	.join( vcf_loc_pair1 )
	.combine(bel_spec1_ch)
	.combine(bel_spec2_ch)
	.filter{ it[5] < it[7] }
	.map{ it[0,1,2,3,4,6,8]}
hon_pairs_ch = Channel.from( "hon" )
	.join( vcf_loc_pair2 )
	.combine(hon_spec1_ch)
	.combine(hon_spec2_ch)
	.filter{ it[5] < it[7] }
	.map{ it[0,1,2,3,4,6,8]}
pan_pairs_ch = Channel.from( "pan" )
	.join( vcf_loc_pair3 )
	.combine(pan_spec1_ch)
	.combine(pan_spec2_ch)
	.filter{ it[5] < it[7] }
	.map{ it[0,1,2,3,4,6,8]}
bel_pairs_ch.concat( hon_pairs_ch, pan_pairs_ch  ).set { all_fst_pairs_ch }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.9
// run fst on actual populations
process fst_run {
	label 'L_32g1h_fst_run'

	input:
	set val( loc ), file( vcf ), file( vcfidx ), file( pop ), file( outlier_tab ), val( spec1 ), val( spec2 ), val( subset_type ) from all_fst_pairs_ch.combine( subset_type_ch )

	output:
	set val( "${spec1}${loc}-${spec2}${loc}_${subset_type}" ), file( "*_random_fst_a00.tsv" ) into rand_header_ch
	set val( "${spec1}${loc}-${spec2}${loc}_${subset_type}" ), val( loc ), val( spec1 ), val( spec2 ), file( "${loc}.${subset_type}.vcf.gz" ), file( "col1.pop" ), file( "prep.pop" ) into rand_body_ch

	script:
	"""
	if [ "${subset_type}" == "subset_non_diverged" ];then
		awk -v OFS="\\t" '{print \$2,\$3,\$4}' ${outlier_tab} > diverged_regions.bed 
		SUBSET="--exclude-bed diverged_regions.bed"
	else
		SUBSET=""
	fi

	vcftools --gzvcf ${vcf[0]} \
		\$SUBSET \
		--recode \
		--stdout | bgzip > ${loc}.${subset_type}.vcf.gz

	tabix ${loc}.${subset_type}.vcf.gz

	echo -e "0000\treal_pop" > idx.txt

	vcfsamplenames ${loc}.${subset_type}.vcf.gz | \
		awk '{print \$1"\\t"substr(\$1, length(\$1)-5, length(\$1))}'  > prep.pop
	grep ${spec1} ${pop} > pop1.txt
	grep ${spec2} ${pop} > pop2.txt
	
	vcftools --gzvcf ${loc}.${subset_type}.vcf.gz \
		--weir-fst-pop pop1.txt \
		--weir-fst-pop pop2.txt \
		--stdout 2> fst.log 1> tmp.txt

	grep "^Weir" fst.log | sed 's/.* //' | paste - - > fst.tsv
	echo -e "idx\\ttype\\tmean_fst\\tweighted_fst" > ${spec1}${loc}-${spec2}${loc}_${subset_type}_random_fst_a00.tsv
	paste idx.txt fst.tsv >> ${spec1}${loc}-${spec2}${loc}_${subset_type}_random_fst_a00.tsv

	rm fst.tsv fst.log pop1.txt pop2.txt tmp.txt idx.txt

	awk '{print \$1}' prep.pop > col1.pop
	"""
}
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.10
// create indexes for permutation itteration
Channel
	.from( ('0'..'9'))
	.map{ "0" + it } //.into{ sub_pre_ch; sub_pre_ch2 }
	.into{ singles_ch; tens_ch }

singles_ch
	.combine(tens_ch)
	.map{ it[0]+it[1] }
	.toSortedList()
	.flatten()
	.into{ sub_pre_ch; sub_pre_ch2 }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.11
// for each itteration run fst on 100
// permutations of population assignment
process random_bodies {
	label 'L_32g6h_fst_run'

	input:
	set val( run ), val( loc ), val( spec1 ), val( spec2 ), file( vcf ), file( col1 ), file( prepop ), val( pre ) from rand_body_ch.combine(sub_pre_ch)

	output:
	set val( run ), file("*_random_fst_b${pre}.tsv") into rand_body_out_ch

	script:
	"""
	for k in {00..99}; do
	echo "Iteration_"\$k
	echo -e "${pre}\$k\trandom" > idx.txt

	awk '{print \$2}' ${prepop} | shuf > col2.pop # premutation happens here
	paste ${col1} col2.pop > rand.pop

	grep "${spec1}${loc}\$" rand.pop > r_pop1.pop
	grep "${spec2}${loc}\$" rand.pop > r_pop2.pop

	vcftools --gzvcf ${vcf} \
		--weir-fst-pop r_pop1.pop \
		--weir-fst-pop r_pop2.pop \
		--stdout  2> fst.log 1> tmp.txt

	grep "^Weir" fst.log | sed 's/.* //' | paste - - > fst.tsv
	paste idx.txt fst.tsv >> ${run}_random_fst_b${pre}.tsv

	rm fst.tsv fst.log rand.pop col2.pop r_pop1.pop r_pop2.pop tmp.txt 
	done
	"""
}
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.12
// collect all itterations and compile
// output for each population pair
process compile_random_results {
	label 'L_20g2h_compile_rand'
	publishDir "../../2_analysis/fst_signif/random", mode: 'copy' 

	input:
	set val( run ), file( body ), file( head ) from rand_body_out_ch.groupTuple().join(rand_header_ch, remainder: true)

	output:
	file("${run}_random_fst.tsv.gz") into random_lists_result

	script:
	"""
	cat ${head} > ${run}_random_fst.tsv
	cat ${body} >> ${run}_random_fst.tsv
	gzip ${run}_random_fst.tsv
	"""
}
```


```{r , eval = FALSE, hilang = 'nf'}
// -----------------------------------------
// repeat the same procedure for adaptation
// (permuting location within species)

// git 12.13
// prepare species channel
Channel
	.from( "nig", "pue", "uni")
	.set{ species_ch }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.14
// define location set
Channel.from( [[1, "bel"], [2, "hon"], [3, "pan"]]).into{ locations_ch_1;locations_ch_2 }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.15
// create location pairs
locations_ch_1
	.combine(locations_ch_2)
	.filter{ it[0] < it[2] }
	.map{ it[1,3]}
	.combine( species_ch )
	.combine( vcf_adapt )
	.combine( outlier_tab2 )
	.combine( subset_type_ch2 )
	.set{ vcf_location_combo_adapt }
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.16
// collapsed analog to git 12.6 & 9
// subset vcf by species and
// run fst on actual populations
process fst_run_adapt {
	label 'L_32g1h_fst_run'

	input:
	set val( loc1 ), val( loc2 ), val( spec ), val( vcf_indx) , file( vcf ), file( outlier_tab ), val( subset_type ) from vcf_location_combo_adapt

	output:
	set val( "${spec}${loc1}-${spec}${loc2}_${subset_type}" ), file( "*_random_fst_a00.tsv" ) into rand_header_adapt_ch
	set val( "${spec}${loc1}-${spec}${loc2}_${subset_type}" ), val( spec ), val( loc1 ), val( loc2 ), file( "${spec}.${subset_type}.vcf.gz" ), file( "col1.pop" ), file( "prep.pop" ) into rand_body_adapt_ch

	script:
	"""
	vcfsamplenames ${vcf[0]} | \
		grep ${spec} > ${spec}.pop

	if [ "${subset_type}" == "subset_non_diverged" ];then
		awk -v OFS="\\t" '{print \$2,\$3,\$4}' ${outlier_tab} > diverged_regions.bed 
		SUBSET="--exclude-bed diverged_regions.bed"
	else
		SUBSET=""
	fi

	vcftools --gzvcf ${vcf[0]} \
		\$SUBSET \
		--keep ${spec}.pop \
		--mac 3 \
		--recode \
		--stdout | bgzip > ${spec}.${subset_type}.vcf.gz

	tabix ${spec}.${subset_type}.vcf.gz

	echo -e "0000\treal_pop" > idx.txt

	vcfsamplenames ${spec}.${subset_type}.vcf.gz | \
		awk '{print \$1"\\t"substr(\$1, length(\$1)-5, length(\$1))}'  > prep.pop
	grep ${loc1} ${spec}.pop > pop1.txt
	grep ${loc2} ${spec}.pop > pop2.txt
	
	vcftools --gzvcf ${spec}.${subset_type}.vcf.gz \
		--weir-fst-pop pop1.txt \
		--weir-fst-pop pop2.txt \
		--stdout 2> fst.log 1> tmp.txt

	grep "^Weir" fst.log | sed 's/.* //' | paste - - > fst.tsv
	echo -e "idx\\ttype\\tmean_fst\\tweighted_fst" > ${spec}${loc1}-${spec}${loc2}_${subset_type}_random_fst_a00.tsv
	paste idx.txt fst.tsv >> ${spec}${loc1}-${spec}${loc2}_${subset_type}_random_fst_a00.tsv

	rm fst.tsv fst.log pop1.txt pop2.txt tmp.txt idx.txt

	awk '{print \$1}' prep.pop > col1.pop
	"""
}
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.17
// for each itteration run fst on 100
// permutations of location assignment
process random_bodies_adapt {
	label 'L_32g6h_fst_run'

	input:
	set val( run ), val( spec ), val( loc1 ), val( loc2 ), file( vcf ), file( col1 ), file( prepop ), val( pre ) from rand_body_adapt_ch.combine(sub_pre_ch2)

	output:
	set val( run ), file("*_random_fst_b${pre}.tsv") into rand_body_out_adapt_ch

	script:
	"""
	for k in {00..99}; do
	echo "Iteration_"\$k
	echo -e "${pre}\$k\trandom" > idx.txt

	awk '{print \$2}' ${prepop} | shuf > col2.pop # premutation happens here
	paste ${col1} col2.pop > rand.pop

	grep "${spec}${loc1}\$" rand.pop > r_pop1.pop
	grep "${spec}${loc2}\$" rand.pop > r_pop2.pop

	vcftools --gzvcf ${vcf} \
		--weir-fst-pop r_pop1.pop \
		--weir-fst-pop r_pop2.pop \
		--stdout  2> fst.log 1> tmp.txt

	grep "^Weir" fst.log | sed 's/.* //' | paste - - > fst.tsv
	paste idx.txt fst.tsv >> ${run}_random_fst_b${pre}.tsv

	rm fst.tsv fst.log rand.pop col2.pop r_pop1.pop r_pop2.pop tmp.txt 
	done
	"""
}
```


```{r , eval = FALSE, hilang = 'nf'}
// git 12.18
// collect all itterations and compile
// output for each location pair
process compile_random_results_adapt {
	label 'L_20g2h_compile_rand'
	publishDir "../../2_analysis/fst_signif/random/adapt", mode: 'copy' 

	input:
	set val( run ), file( body ), file( head ) from rand_body_out_adapt_ch.groupTuple().join(rand_header_adapt_ch, remainder: true)

	output:
	file("${run}_random_fst.tsv.gz") into random_lists_adapt_result

	script:
	"""
	cat ${head} > ${run}_random_fst.tsv
	cat ${body} >> ${run}_random_fst.tsv
	gzip ${run}_random_fst.tsv
	"""
}
```
:::

---
