---
output: html_document
editor_options:
  chunk_output_type: console
css: highlight.css
---

```{r setup, include=FALSE}
knitr::knit_hooks$set(source = function(x, options) {
  if (!is.null(options$hilang)) {
      code_open <- "\n\n<div class=\"sourceCode\">\n<pre class=\"sourceCode\">\n<code class=\"sourceCode\">"
      code_close <- "\n</code>\n</pre>\n</div>\n"
      code_body <- highr::hi_andre(x, language = options$hilang, format = "html")
    stringr::str_c(
      code_open,
      knitr:::indent_block(paste(code_body, collapse = '\n'), ""),
      code_close
    )
  } else {
    stringr::str_c("\n\n```", tolower(options$engine), "\n",
                   paste(x, collapse = '\n'), "\n```\n\n")

  }
})
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')
```


```{r, include=FALSE}
# source('R/draw_workflow.R')
# prod_het <- tibbler(c('inds_out', 'win_out'))
```

# (git 10) Analysis VIII (admixture)

## Summary

The population recombination rate is estimated within the [**nextflow**](https://www.nextflow.io/) script `analysis_XX.nf` (located under `$BASE_DIR/nf/10_analysis_admixture/`), which runs on the XX data set.
Below is an overview of the steps involved in the analysis.
(The <span style="color:#4DAF4A">green dot</span> indicates the genotype input, <span style="color:#E41A1C">red arrows</span> depict output that is exported for further use.)

<div style="max-width:500px; margin:auto;">
```{r, echo = FALSE, warning = FALSE, message = FALSE}
# girafe( ggobj = dot_plot(file = 'docs/analysis_het.dot', git = 7, point_types = prod_het),
#        width_svg = 6, height_svg = 6)
```
</div>

## Details of `analysis_xx.nf`

### Data preparation

The nextflow script starts by opening the genotype data.

<div class="kclass">
```{r , eval = FALSE, hilang = 'nf'}
#!/usr/bin/env nextflow
// git 10.1
// open genotype data
Channel
	.fromFilePairs("../../1_genotyping/4_phased/phased_mac2.vcf.{gz,gz.tbi}")
	.set{ vcf_ch }

// git 10.2
// Set different k values for the admixture analysis
Channel
	.from( 2..15 )
	.set{ k_ch }

// git 10.3
// load Fst outlier regions
Channel
	.fromPath("../../ressources/plugin/poptrees/outlier.bed")
	.splitCsv(header:true, sep:"\t")
	.map{ row -> [ chrom:row.chrom, start:row.start, end:row.end, gid:row.gid ] }
	.combine( vcf_ch )
	.set{ vcf_admx }

// git 10.4
// subset genotypes to the outlier region and reformat
process plink12 {
	label 'L_20g2h_plink12'
	tag "${grouping.gid}"

	input:
	set val( grouping ),  val( vcfidx ), file( vcf ) from vcf_admx

	output:
	set val( grouping ), file( "hapmap.*.ped" ), file( "hapmap.*.map" ), file( "hapmap.*.nosex" ), file( "pop.txt" ) into admx_plink

	script:
	"""
	echo -e "CHROM\\tSTART\\tEND" > outl.bed
	echo -e "${grouping.chrom}\\t${grouping.start}\\t${grouping.end}" >> outl.bed

	vcfsamplenames ${vcf[0]} | \
		grep -v "tor\\|tab\\|flo" | \
		awk '{print \$1"\\t"\$1}' | \
		sed 's/\\t.*\\(...\\)\\(...\\)\$/\\t\\1\\t\\2/g' > pop.txt

	vcftools \
		--gzvcf ${vcf[0]} \
		--keep pop.txt \
		--bed outl.bed \
		--plink \
		--out admx_plink

	plink \
		--file admx_plink \
		--recode12 \
		--out hapmap.${grouping.gid}
	"""
}

// git 10.5
// combine genoutype subsets with k values
admx_prep  = k_ch.combine( admx_plink )

// git 10.6
// run admixture
process admixture_all {
	label 'L_20g4h_admixture_all'
	publishDir "../../2_analysis/admixture/", mode: 'copy'
	tag "${grouping.gid}.${k}"

	input:
	set  val( k ), val( grouping ), file( ped ), file( map ), file( nosex ), file( pop ) from admx_prep

	output:
	set val( "dummy" ), file( "*.out" ), file( "*.Q" ), file( "*.txt" ) into admx_log

	script:
	"""
	mv ${pop} pop.${grouping.gid}.${k}.txt
	admixture --cv ${ped} ${k} | tee log.${grouping.gid}.${k}.out
	"""
}
```
</div>

Finally, we are done with XX.

---
