---
output: html_document
editor_options:
  chunk_output_type: console
---
# Figure --xx--

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      comment = "#>",
                      dev = "ragg_png",
                      eval = FALSE)
knitr::opts_knit$set(root.dir = '../')

library(ragg)
```

## Summary

This is the accessory documentation of Figure --xx--.

The Figure can be recreated by running the **R** script `plot_F--xx--.R` from a (`bash` terminal):

```sh
cd $BASE_DIR

Rscript --vanilla R/fig/plot_F--xx--.R 2_analysis/dxy/50k/ 2_analysis/fst/50k/

```

## Details of `plot_F--xx--.R`

In the following, the individual steps of the R script are documented.
It is an executable R script that depends on the accessory R package [**GenomicOriginsScripts**](https://k-hench.github.io/GenomicOriginsScripts), [**BAMMtools**](https://cran.r-project.org/web/packages/BAMMtools/) and on the package [**hypoimg**](https://k-hench.github.io/hypoimg).

### Config

The scripts start with a header that contains copy & paste templates to execute interactively or debug the script:

```{r , eval = FALSE}

```

The next section processes the input from the command line.
It stores the arguments in the vector `args`.
The needed R packages are loaded and the script name and the current working directory are stored inside variables (`script_name`, `plot_comment`).
This information will later be written into the meta data of the figure to help us tracing back the scripts that created the figures in the future.

Then we drop all the imported information besides the arguments following the script name and print the information to the terminal.

```{r, eval = FALSE}

```

```r
#> ── Script: R/fig/plot_F--xx--.R ──────────────────────────────────────────────
#> Parameters read:
#>  ★ 1:
#>  ★ 2:
#>  ★ 3:
#> ─────────────────────────────────────────── /current/working/directory ──
```
The directories for the different data types are received and stored in respective variables.
Also, we set a few parameters for the plot layout:

```{r}
# config -----------------------
... <- as.character(args[1])

```

### Actual Script Start

```{r, message = FALSE, warning = FALSE}
#!/usr/bin/env Rscript
#
# Context: this script depends on the input file 2_analysis/summaries/fst_permutation_summary.tsv
#          which is created by R/fst_permutation.R
#
# run from terminal:
# Rscript --vanilla R/fig/plot_F1.R \
#   2_analysis/fst/50k/ \
#   2_analysis/summaries/fst_globals.txt \
#   2_analysis/summaries/fst_permutation_summary.tsv \
#   2_analysis/fotl/concat_R24ed.treefile
# ===============================================================
# This script produces Figure 1 of the study "Rapid radiation in a highly
# diverse marine environment" by Hench, Helmkampf, McMillan and Puebla
# ---------------------------------------------------------------
# ===============================================================
# args <- c('2_analysis/fst/50k/', '2_analysis/summaries/fst_globals.txt',
#           '2_analysis/summaries/fst_permutation_summary.tsv',
#           "2_analysis/fotl/concat_R24ed.treefile")
# script_name <- "R/fig/plot_F1.R"
args <- commandArgs(trailingOnly = FALSE)
# setup -----------------------
renv::activate()
library(GenomicOriginsScripts)
library(hypoimg)
library(hypogen)
library(patchwork)
library(ape)
library(ggraph)
library(tidygraph)
library(stringr)
library(ggtree)

cat('\n')
script_name <- args[5] %>%
  str_remove(., '--file=')

plot_comment <- script_name %>%
  str_c('mother-script = ', getwd(), '/', .)

args <- process_input(script_name, args)
# config -----------------------
fst_dir <- as.character(args[1])
fst_globals <- as.character(args[2])
fst_permutation_file <- as.character(args[3])
tree_file <- as.character(args[4])
wdh <- .3          # The width of the boxplots
scaler <- 20       # the ratio of the Fst and the dxy axis (legacy - not really needed anymore)
clr_sec <- 'gray'  # the color of the secondary axis (dxy)
# start script -------------------

tree <- read.tree(tree_file)
tree_rooted <- root(phy = tree, outgroup = "Epinephelus_maculatus")
clr_neutral <- rgb(.2, .2, .2)
clr_highlight <- "gray40" #"#FF8029"
clr_stars <- "firebrick"

### Edit tip labels
tree_rooted$tip.label <- tree_rooted$tip.label %>%
  str_replace(pattern = "20864abehon", "Hypoplectrus_aberrans") %>%
  str_replace(pattern = "20642gumhon", "Hypoplectrus_gummigutta") %>%
  str_replace(pattern = "18238indbel", "Hypoplectrus_indigo") %>%
  str_replace(pattern = "PL17_122maybel", "Hypoplectrus_maya") %>%
  str_replace(pattern = "18906nigpan", "Hypoplectrus_nigricans") %>%
  str_replace(pattern = "18434puepan", "Hypoplectrus_puella") %>%
  str_replace(pattern = "20613ranhon", "Hypoplectrus_randallorum") %>%
  str_replace(pattern = "18448unipan", "Hypoplectrus_unicolor") %>%
  str_replace(pattern = "PL17_160floflo", "Hypoplectrus_floridae") %>%
  str_replace(pattern = "20478tabhon", "Serranus_tabacarius") %>%
  str_replace(pattern = "s_tort_3torpan", "Serranus_tortugarum") %>%
  #
  str_replace(pattern = "([A-Z])([a-z])[a-z]*_([a-z]*)", "\\1\\2. \\3")  %>%
  str_replace(pattern = "Ce.", "Cp.")%>%
  str_replace(pattern = "Za.", "Pl.") %>%
  str_replace(pattern = "Hy.", "H.") %>%
  str_replace(pattern = "Di.", "D.") %>%
  str_replace(pattern = "Ep.", "E.")

### Prepare tree, categorize support values and define group
tree_plus <- ggtree(tree_rooted, layout = "rectangular", ladderize = TRUE, right = TRUE) %>%
  #flip(25,26) %>%
  .$data %>%
  mutate(support = as.numeric(label),
         support_class = cut(support, c(0,50,70,90,100)) %>%
           as.character() %>% factor(levels = c("(0,50]", "(50,70]", "(70,90]", "(90,100]")))

hamlets <- tree_plus$label[grepl(pattern = "H. ", tree_plus$label)]

tree_plus <- tree_plus %>%
  groupOTU(.node = hamlets)

### Define groups
our_taxa <- c("H. aberrans", "H. gummigutta", "H. indigo", "H. maya", "H. nigricans", "H. puella",
              "H. randallorum", "H. unicolor", "H. floridae", "Se. tabacarius", "Se. tortugarum")

# hermaphrodites <- tree_plus %>% filter(node %in% c(60, 65) | label == "Se. hepatus") %>%
#   mutate(x = x - branch.length / 2, star = "\U2605")

hermaphrodites <- tree_plus %>% filter(node %in% c(60, 63)) %>%
  mutate(x = if_else(node == 60, x - branch.length, x - branch.length *.5),
         y = if_else(node == 60, .5 * (y + tree_plus$y[tree_plus$node == 61]), y),
         star = "\U2605")

c1 <- "transparent"
# c2 <- rgb(0, 0, 0, .1)
# c3 <- rgb(0, 0, 0, .2)
c2 <- prismatic::clr_alpha(clr_highlight, .1)
c3 <- prismatic::clr_alpha(clr_highlight, .2)

grad_mat <- c(c1, c2, c3, c3)
dim(grad_mat) <- c(1, length(grad_mat))
grob_grad <- rasterGrob(grad_mat,
                        width = unit(1, "npc"),
                        height = unit(1, "npc"),
                        interpolate = TRUE)

blank_hamlet <- hypoimg::hypo_outline %>%
  ggplot()+
  coord_equal()+
  geom_polygon(aes(x, y), color = rgb(0,0,0,.3), fill = rgb(1, 1, 1, .3), size = .1)+
  theme_void()

### Draw tree
(p_tree <- ggtree(tree_plus,
                 aes(color = group), size = .2) +
  annotation_custom(grob = grob_grad,
                    ymin = .2, ymax = 12.5,
                    xmin = .04, xmax = .125)+
  annotation_custom(grob = ggplotGrob(blank_hamlet),
                    xmin = 0.09, xmax = .125,
                    ymin = 1.2, ymax = 10.5) +
  geom_tiplab(aes(color = group, label = if_else(label %in% our_taxa, str_c(label,"*"), label)),   # add asterisks to our taxa
              size = 1.3, hjust = -.1,
              fontface = 'italic') +
  ggplot2::xlim(0, 0.12) +   # add extra space for long labels
  geom_nodepoint(data = tree_plus %>% filter(!is.na(support_class)),
                   aes(fill = support_class,
                     size = support_class),
                 shape = 21, color = clr_neutral) +
  geom_text(data = hermaphrodites, aes(x = x, y = y, label = star),
            family = "DejaVu Sans", color = clr_stars,
            size = 2, vjust = .35) +
  scale_color_manual(values = c(clr_neutral, clr_neutral, clr_highlight)) +
  scale_fill_manual(values = c(`(0,50]`   = "transparent",
                               `(50,70]`  = "white",
                               `(70,90]`  = "gray",
                               `(90,100]` = "black"),
                    drop = FALSE) +
  scale_size_manual(values = c(`(0,50]`   = 0,
                               `(50,70]`  = .8,
                               `(70,90]`  = .8,
                               `(90,100]` = .8),
                    na.value = 0,
                    drop = FALSE) +
  geom_treescale(color = clr_neutral,
                 fontsize = 2, linesize = .2,
                 x = 0.045, y = 1.5) +
  guides(fill = guide_legend(title = "Node Support Class", title.position = "top", override.aes = list(color = clr_neutral), nrow = 2),
         size = guide_legend(title = "Node Support Class", title.position = "top", override.aes = list(color = clr_neutral), nrow = 2),
         color = 'none') +
    coord_cartesian(xlim = c(-.005,.125), ylim = c(0,45), expand = 0) +
    theme_void() +
  theme(legend.position = c(0.05,0),
        legend.justification = c(0,0),
        legend.title.align = 0,
        legend.key.height = unit(8,"pt"),
        legend.key.width = unit(6,"pt"),
        legend.text = element_text(color = clr_neutral),
        legend.title = element_text(color = clr_neutral))
)

globals <- vroom::vroom(fst_globals, delim = '\t',
                        col_names = c('loc','run','mean','weighted')) %>%
  mutate(run = str_c(str_sub(run,1,3),loc,'-',str_sub(run,5,7),loc),
         run = fct_reorder(run,weighted))

# sort run by average genome wide Fst
run_ord <- tibble(run = levels(globals$run),
                  run_ord = seq_along(levels(globals$run)))

# load fst permutation results
fst_sig_attach <- read_tsv(fst_permutation_file) %>%
  filter( subset_type == "whg" ) %>%
  mutate(loc = str_sub(run, -3, -1)) %>%
  group_by(loc) %>%
  mutate(loc_n = 28,#length(loc),
         fdr_correction_factor =  sum(1 / 1:loc_n),
         fdr_alpha = .05 / fdr_correction_factor,
         is_sig = p_perm > fdr_alpha) %>%
  ungroup()

# create network annotation
# underlying structure for the network plots
networx <- tibble( loc = c('bel','hon', 'pan'),
                   n = c(5, 6, 3),
                   label = list(str_c(c('ind','may','nig','pue','uni'),'bel'),
                                str_c(c('abe','gum','nig','pue','ran','uni'),'hon'),
                                str_c(c('nig','pue','uni'),'pan')),
                   weight = c(1,1.45,1)) %>%
  purrr::pmap_dfr(network_layout) %>%
  mutate(edges = map(edges, function(x){x %>% left_join(globals,by = "run") }))

# plot the individual networks by location
plot_list <- networx %>%
  purrr::pmap(plot_network, node_lab_shift = .2)

# combine the networks into a single grob
p_net <- cowplot::plot_grid(
  plot_list[[1]] + theme(legend.position = "none"),
  plot_list[[2]] + theme(legend.position = "none"),
  plot_list[[3]] + theme(legend.position = "none"),
  ncol = 3) %>%
  cowplot::as_grob()

p2 <- globals %>%
  left_join(fst_sig_attach %>% mutate(run = factor(run, levels = levels(globals$run))) ) %>%
  ggplot(aes(color = loc)) +
  geom_bar(aes(x = as.numeric(run),
                   y = weighted,
               alpha = is_sig,
               fill = after_scale(clr_lighten(color))),
               stat = "identity",size = .2, width = .8)+
  annotation_custom(p_net, ymin = .05, xmax = 24.5) +
  scale_x_continuous(name = "Pair of sympatric species",
                     breaks = 1:28) +
  scale_y_continuous(name = expression(italic(F[ST])))+
  scale_color_manual(values = c(make_faint_clr('bel'),
                                make_faint_clr('hon'),
                                make_faint_clr('pan'))[c(2, 4, 6)])+
  scale_shape_manual(values = c(`TRUE` = 1, `FALSE` = 21)) +
  scale_alpha_manual(values = c(`TRUE` = .15, `FALSE` = 1)) +
  coord_cartesian(xlim = c(0,29),
                  expand = c(0,0))+
  theme_minimal()+
  theme(text = element_text(size = plot_text_size),
        legend.position = 'none',
        strip.placement = 'outside',
        strip.text = element_text(size = 12),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.y.right = element_text(color = clr_sec),
        axis.title.y.right = element_text(color = clr_sec))
# assemble panel c-e
clr_alt <- clr
clr_alt[["uni"]] <- "lightgray"
pca_fish_scale <- 1.15

pca_fish_pos <- tibble(pop = GenomicOriginsScripts::pop_levels,
                       short = str_sub(pop, 1, 3),
                       loc = str_sub(pop, 4, 6),
                       width = c(bel = .08, hon =.08, pan = .09)[loc] * pca_fish_scale,
                       height = c(bel = .08, hon =.08, pan = .09)[loc] * pca_fish_scale) %>%
  arrange(loc) %>%
  mutate(x = c(-.18, -.01, .03, -.03, .075,
               -.04, -.2, -.02, -.01, -.02, -.01,
               -.15, 0, .05),
         y = c(.02, .27, -.13, -.03, .05,
               -.1, .05, 0, .075, -.23, .2,
               .06, -.2, .2)) %>%
  select(-pop) %>%
  group_by(loc) %>%
  nest()

pcas <- c("bel", "hon", "pan") %>% map(pca_plot)

fish_tib <- tibble(short = names(clr)[!names(clr) %in% c("flo", "tab", "tor")],
       x = c(0.5,  3.5,  7,  9.7, 12.25, 15.25, 18, 21.5)
       )
# ===================
key_sz <- .75
p_leg <- fish_tib %>%
  ggplot() +
  coord_equal(xlim = c(-.05, 24), expand = 0) +
  geom_tile(aes(x = x, y = 0,
                fill = short,
                color = after_scale(prismatic::clr_darken(fill, .25))),
            width = key_sz, height = key_sz, size = .3) +
  geom_text(aes(x = x + .6, y = 0,
                label = str_c("H. ", sp_names[short])),
            hjust = 0, fontface = "italic", size = plot_text_size / ggplot2:::.pt) +
  pmap(fish_tib, plot_fish_lwd, width = 1, height = 1, y = 0) +
  scale_fill_manual(values = clr, guide = FALSE) +
  theme_void()

p_combined <- ((wrap_elements(plot = p_tree +
                                theme(axis.title = element_blank(),
                                      text = element_text(size = plot_text_size)),
                              clip = FALSE) + p2 ) /
                 (pcas %>% wrap_plots()) +
                 plot_layout(heights = c(1,.75)) +
                 plot_annotation(tag_levels = 'a') &
                 theme(text = element_text(size = plot_text_size),
                       plot.background = element_rect(fill = "transparent",
                                                      color = "transparent")))

p_done <- cowplot::plot_grid(p_combined, p_leg, ncol = 1, rel_heights = c(1,.06))

scl <- .75
hypo_save(p_done, filename = 'figures/F1.pdf',
          width = 9 * scl,
          height = 6 * scl,
          device = cairo_pdf,
          bg = "transparent",
          comment = plot_comment)
```

Finally, we can export Figure --xx--.

```{r, eval = FALSE}
scl <- .75
hypo_save(p_done, filename = 'figures/F--xx--.pdf',
          width = 9 * scl,
          height = 6 * scl,
          device = cairo_pdf,
          bg = "transparent",
          comment = plot_comment)
```

---
