---
output: html_document
editor_options:
  chunk_output_type: console
---
# Supplementary Figure 4

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      comment = "#>",
                      dev = "ragg_png",
                      eval = FALSE)
knitr::opts_knit$set(root.dir = '../')

library(ragg)
```


```{r , include = FALSE}
args <- c('2_analysis/fst/50k/',
          '2_analysis/summaries/fst_outliers_998.tsv',
          '2_analysis/summaries/fst_globals.txt')

library(GenomicOriginsScripts)
library(hypoimg)
library(hypogen)
library(vroom)
library(ggtext)

data_path <- as.character(args[1])
outlier_file <- as.character(args[2])
globals_file <- as.character(args[3])
```

## Summary

This is the accessory documentation of Figure S4.
The Figure can be recreated by running the **R** script `plot_SF4.R`:

```sh
cd $BASE_DIR

Rscript --vanilla R/fig/plot_SF4.R \
    2_analysis/fst/50k/ \
    2_analysis/summaries/fst_outliers_998.tsv \
    2_analysis/summaries/fst_globals.txt
```

## Details of `plot_SF4.R`

In the following, the individual steps of the R script are documented.
It is an executable R script that depends on the accessory R package [**GenomicOriginsScripts**](https://k-hench.github.io/GenomicOriginsScripts), as well as on the packages [**hypoimg**](https://k-hench.github.io/hypoimg), [**hypogen**](https://k-hench.github.io/hypogen) and [**patchwork**](https://patchwork.data-imaginist.com/)

### Config

The scripts start with a header that contains copy & paste templates to execute or debug the script:

```{r, eval = FALSE}
#!/usr/bin/env Rscript
# run from terminal:
# Rscript --vanilla R/fig/plot_SF4.R \
#     2_analysis/fst/50k/ \
#     2_analysis/summaries/fst_outliers_998.tsv \
#     2_analysis/summaries/fst_globals.txt
# ===============================================================
# This script produces Suppl. Figure 4 of the study "Rapid radiation in a
# highly diverse marine environment" by Hench, Helmkampf, McMillan and Puebla
# ---------------------------------------------------------------
# ===============================================================
# args <- c('2_analysis/fst/50k/',
#           '2_analysis/summaries/fst_outliers_998.tsv',
#           '2_analysis/summaries/fst_globals.txt')
# script_name <- "R/fig/plot_SF4.R"
args <- commandArgs(trailingOnly = FALSE)
```

The next section processes the input from the command line.
It stores the arguments in the vector `args`.
The needed R packages are loaded and the script name and the current working directory are stored inside variables (`script_name`, `plot_comment`).
This information will later be written into the meta data of the figure to help us tracing back the scripts that created the figures in the future.

Then we drop all the imported information besides the arguments following the script name and print the information to the terminal.

```{r, eval = FALSE}
# setup -----------------------
renv::activate()
library(GenomicOriginsScripts)
library(hypoimg)
library(hypogen)
library(vroom)
library(ggtext)

cat('\n')
script_name <- args[5] %>%
  str_remove(.,'--file=')

plot_comment <- script_name %>%
  str_c('mother-script = ',getwd(),'/',.)

args <- process_input(script_name, args)
```

```r
#> ── Script: R/fig/plot_SF4.R ────────────────────────────────────────────
#> Parameters read:
#> ★ 1: 2_analysis/fst/50k/
#> ★ 2: 2_analysis/summaries/fst_outliers_998.tsv
#> ★ 3: 2_analysis/summaries/fst_globals.txt
#> ────────────────────────────────────────── /current/working/directory ──
```

The directory containing the PCA data is received and stored in a variable.
Also the default color scheme is updated and the size of the hamlet ann.

```{r, message = FALSE, warning = FALSE}
# config -----------------------
data_path <- as.character(args[1])
outlier_file <- as.character(args[2])
globals_file <- as.character(args[3])
```


```{r, message = FALSE, warning = FALSE}
# load data -------------------
# locate fst data files
files <- dir(data_path,pattern = '.50k.windowed.weir.fst.gz')
```


```{r, message = FALSE, warning = FALSE}
# extract run names from data file names
run_files <- files %>%
  str_sub(.,1,11) %>%
  str_replace(.,pattern = '([a-z]{3})-([a-z]{3})-([a-z]{3})', '\\2\\1-\\3\\1')
```


```{r, message = FALSE, warning = FALSE}
# load genome wide average fst values for each run
globals <- vroom::vroom(globals_file, delim = '\t',
                        col_names = c('loc','run','mean','weighted')) %>%
  separate(run, into = c('pop1','pop2')) %>%
  mutate(run = str_c(pop1,loc,'-',pop2,loc),
         run = fct_reorder(run,weighted))
```


```{r, message = FALSE, warning = FALSE}
# load all windowed fst data and collapse in to a single data frame
data <- purrr::pmap(tibble(file = str_c(data_path,files),
                           run = run_files),
                    hypo_import_windows) %>%
  bind_rows() %>%
  purrr::set_names(., nm = c('CHROM', 'BIN_START', 'BIN_END', 'N_VARIANTS',
                      'WEIGHTED_FST', 'MEAN_FST', 'GSTART', 'POS', 'GPOS', 'run')) %>%
  mutate(pop1 = str_sub(run,1,3),
         pop2 = str_sub(run,8,10),
         loc = str_sub(run,4,6),
         run_label = str_c("*H. ", sp_names[pop1],"* - *H. ", sp_names[pop2],"*<br>(",loc_names[loc],")" ))
```


```{r, message = FALSE, warning = FALSE}
# create table for the indication of genome wide average fst in the plot background
# (rescale covered fst range to the extent of the genome)
global_bar <- globals %>%
  select(weighted,run) %>%
  mutate(run = as.character(run)) %>%
  setNames(.,nm = c('fst','run')) %>%
  pmap(.,fst_bar_row_run) %>%
  bind_rows() %>%
  mutate(pop1 = str_sub(run,1,3),
         pop2 = str_sub(run,8,10),
         loc = str_sub(run,4,6),
         run_label = str_c("*H. ", sp_names[pop1],"* - *H. ", sp_names[pop2],"*<br>(",loc_names[loc],")" ),
         run_label = fct_reorder(run_label,xmax_org))
```


```{r, message = FALSE, warning = FALSE}
# pre-calculate secondary x-axis breaks
sc_ax <- scales::cbreaks(c(0,max(globals$weighted)),
                         scales::pretty_breaks(4))
```


```{r, message = FALSE, warning = FALSE}
# compose final figure
p_done <- ggplot()+
  # general plot structure separated by run
  facet_grid( run_label ~ ., as.table = TRUE) +
  # add genome wide average fst in the background
  geom_rect(data = global_bar %>%
              mutate(xmax = xmax * hypo_karyotype$GEND[24]),
            aes(xmin = 0, xmax = xmax,
                ymin = -Inf, ymax = Inf),
            color = rgb(1,1,1,0),
            fill = clr_below) +
  # add LG borders
  geom_vline(data = hypogen::hypo_karyotype,
             aes(xintercept = GEND),
             color = hypo_clr_lg) +
  # add fst data points
  geom_point(data = data  %>%
                      mutate(run_label = factor(run_label,
                                                levels = levels(global_bar$run_label))),
              aes(x = GPOS, y = WEIGHTED_FST),
              size=.2,color = plot_clr) +
  # axis layout
  scale_x_hypo_LG(sec.axis =  sec_axis(~ ./hypo_karyotype$GEND[24],
                                       breaks = (sc_ax$breaks/max(globals$weighted)),
                                       labels = sprintf("%.2f", sc_ax$breaks),
                                       name = expression(Genomic~position/~Genome~wide~weighted~italic(F[ST])))) +
  scale_y_continuous(name = expression(italic('F'[ST])),
                     limits = c(-.1,1),
                     breaks = c(0,.5,1)) +
  # general plot layout
  theme_hypo() +
  theme(strip.text.y = element_markdown(angle = 0),
        strip.background = element_blank(),
        legend.position = 'none',
        axis.title.x = element_text(),
        axis.text.x.bottom = element_text(colour = 'darkgray'))
```

Finally, we can export Figure S4.

```{r, eval = FALSE}
# export final figure
hypo_save(filename = 'figures/SF4.png',
          plot = p_done,
          width = 8,
          height = 12,
          dpi = 600,
          type = "cairo",
          comment = plot_comment)

system("convert figures/SF4.png figures/SF4.pdf")
system("rm figures/SF4.png")
create_metadata <- str_c("exiftool -overwrite_original -Description=\"", plot_comment, "\" figures/SF4.pdf")
system(create_metadata)
```
