---
output: html_document
editor_options:
  chunk_output_type: console
---
# Supplementary Figure 5

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
knitr::opts_knit$set(root.dir = '../')
library(GenomicOriginsScripts)
library(hypoimg)
library(hypogen)
library(vroom)
library(ggtext)

args <- c('2_analysis/fst/50k/',
          '2_analysis/summaries/fst_outliers_998.tsv',
          '2_analysis/summaries/fst_globals.txt')
```

## Summary

This is the accessory documentation of Figure S5.
The Figure can be recreated by running the **R** script `plot_SF5.R`:

```sh
cd $BASE_DIR

Rscript --vanilla R/fig/plot_SF5.R 2_analysis/fst/50k/ \
  2_analysis/summaries/fst_outliers_998.tsv \
  2_analysis/summaries/fst_globals.txt

```

## Details of `plot_SF5.R`

In the following, the individual steps of the R script are documented.
It is an executable R script that depends on the accessory R package [**GenomicOriginsScripts**](https://k-hench.github.io/GenomicOriginsScripts), as well as on the packages [**ggtext**](https://wilkelab.org/ggtext/), [**hypoimg**](https://k-hench.github.io/hypoimg), [**hypogen**](https://k-hench.github.io/hypogen) and [**vroom**](https://vroom.r-lib.org/).

### Config

The scripts start with a header that contains copy & paste templates to execute or debug the script:

```{r, message = FALSE, warning = FALSE}
#!/usr/bin/env Rscript
# run from terminal:
# Rscript --vanilla R/fig/plot_SF5.R 2_analysis/fst/50k/ \
#   2_analysis/summaries/fst_outliers_998.tsv \
#   2_analysis/summaries/fst_globals.txt
# ===============================================================
# This script produces Suppl. Figure 5 of the study "Ancestral variation,
# hybridization and modularity fuel a marine radiation"
# by Hench, Helmkampf, McMillan and Puebla
# ---------------------------------------------------------------
# ===============================================================
# args <- c('2_analysis/fst/50k/',
#           '2_analysis/summaries/fst_outliers_998.tsv',
#           '2_analysis/summaries/fst_globals.txt')
# script_name <- "R/fig/plot_SF5.R"
```

The next section processes the input from the command line.
It stores the arguments in the vector `args`.
The needed R packages are loaded and the script name and the current working directory are stored inside variables (`script_name`, `plot_comment`).
This information will later be written into the meta data of the figure to help us tracing back the scripts that created the figures in the future.

Then we drop all the imported information besides the arguments following the script name and print the information to the terminal.

```{r, eval = FALSE}
args <- commandArgs(trailingOnly=FALSE)
# setup -----------------------
library(GenomicOriginsScripts)
library(hypoimg)
library(hypogen)
library(vroom)
library(ggtext)

cat('\n')
script_name <- args[5] %>%
  str_remove(.,'--file=')

plot_comment <- script_name %>%
  str_c('mother-script = ',getwd(),'/',.)

args <- process_input(script_name, args)
```

```r
#> ── Script: R/fig/plot_SF5.R ────────────────────────────────────────────
#> Parameters read:
#> ★ 1: 2_analysis/fst/50k/
#> ★ 2: 2_analysis/summaries/fst_outliers_998.tsv
#> ★ 3: 2_analysis/summaries/fst_globals.txt
#> ────────────────────────────────────────── /current/working/directory ──
```

The directory containing the $F_{ST}$ data, and the files containing the locations
of the $F_{ST}$ outlier regions and the genome wide $F_{ST}$ averages are
received and stored in a variable.

```{r, message = FALSE, warning = FALSE}
# config -----------------------
data_path <- as.character(args[1])
outlier_file <- as.character(args[2])
globals_file <- as.character(args[3])
```

The files containing the windowed $F_{ST}$ data are located.

```{r, message = FALSE, warning = FALSE}
# load data -------------------
# locate fst data files
files <- dir(data_path, pattern = '.50k.windowed.weir.fst.gz')
```

Based on these data files, the names of the pair wise species comparisons are created.

```{r, message = FALSE, warning = FALSE}
# extract run names from data file names
run_files <- files %>%
  str_sub(.,1,11) %>%
  str_replace(.,pattern = '([a-z]{3})-([a-z]{3})-([a-z]{3})', '\\2\\1-\\3\\1')
```

Then, the genome wide average $F_{ST}$ values are loaded.

```{r, message = FALSE, warning = FALSE}
# load genome wide average fst values for each run
globals <- vroom::vroom(globals_file, delim = '\t',
                        col_names = c('loc','run','mean','weighted')) %>%
  separate(run, into = c('pop1','pop2')) %>%
  mutate(run = str_c(pop1,loc,'-',pop2,loc),
         run = fct_reorder(run,weighted))
```

Next, the windowed $F_{ST}$ data are imported.

```{r, message = FALSE, warning = FALSE}
# load all windowed fst data and collapse in to a single data frame
data <- purrr::pmap(tibble(file = str_c(data_path,files),
                           run = run_files),
                    hypo_import_windows) %>%
  bind_rows() %>%
  purrr::set_names(., nm = c('CHROM', 'BIN_START', 'BIN_END', 'N_VARIANTS',
                      'WEIGHTED_FST', 'MEAN_FST', 'GSTART', 'POS', 'GPOS', 'run')) %>%
  mutate(pop1 = str_sub(run,1,3),
         pop2 = str_sub(run,8,10),
         loc = str_sub(run,4,6),
         run_label = str_c("*H. ", sp_names[pop1],"* - *H. ", sp_names[pop2],"*<br>(",loc_names[loc],")" ))
```

To be able to indicate the genome wide average $F_{ST}$ in the background of the
figure, the $F_{ST}$ values are scaled to the extent of  the hamlet reference genome.
The rescaled $F_{ST}$ values are compiled into a table for plotting.

```{r, message = FALSE, warning = FALSE}
# create table for the indication of genome wide average fst in the plot background
# (rescale covered fst range to the extent of the genome)
global_bar <- globals %>%
  select(weighted,run) %>%
  mutate(run = as.character(run)) %>%
  setNames(.,nm = c('fst','run')) %>%
  pmap(.,fst_bar_row_run) %>%
  bind_rows() %>%
  mutate(pop1 = str_sub(run,1,3),
         pop2 = str_sub(run,8,10),
         loc = str_sub(run,4,6),
         run_label = str_c("*H. ", sp_names[pop1],"* - *H. ", sp_names[pop2],"*<br>(",loc_names[loc],")" ),
         run_label = fct_reorder(run_label,xmax_org))
```

To indicate the genome wide average $F_{ST}$ on a secondary x-axis, the secondary
x-breakes are pre-computed.

```{r, message = FALSE, warning = FALSE}
# pre-calculate secondary x-axis breaks
sc_ax <- scales::cbreaks(c(0,max(globals$weighted)),
                         scales::pretty_breaks(4))
```

Then, the Supplementary Figure is assembled.

```{r, message = FALSE, warning = FALSE}
# compose final figure
p_done <- ggplot()+
  # general plot structure separated by run
  facet_grid( run_label ~ ., as.table = TRUE) +
  # add genome wide average fst in the background
  geom_rect(data = global_bar %>%
              mutate(xmax = xmax * hypo_karyotype$GEND[24]),
            aes(xmin = 0, xmax = xmax,
                ymin = -Inf, ymax = Inf),
            color = rgb(1,1,1,0),
            fill = clr_below) +
  # add LG borders
  geom_vline(data = hypogen::hypo_karyotype,
             aes(xintercept = GEND),
             color = hypo_clr_lg) +
  # add fst data points
  geom_point(data = data  %>%
                      mutate(run_label = factor(run_label,
                                                levels = levels(global_bar$run_label))),
              aes(x = GPOS, y = WEIGHTED_FST),
              size=.2,color = plot_clr) +
  # axis layout
  scale_x_hypo_LG(sec.axis =  sec_axis(~ ./hypo_karyotype$GEND[24],
                                       breaks = (sc_ax$breaks/max(globals$weighted)),
                                       labels = sprintf("%.2f", sc_ax$breaks),
                                       name = expression(Genomic~position/~Genome~wide~weighted~italic(F[ST])))) +
  scale_y_continuous(name = expression(italic('F'[ST])),
                     limits = c(-.1,1),
                     breaks = c(0,.5,1)) +
  # general plot layout
  theme_hypo() +
  theme(strip.text.y = element_markdown(angle = 0),
        strip.background = element_blank(),
        legend.position = 'none',
        axis.title.x = element_text(),
        axis.text.x.bottom = element_text(colour = 'darkgray'))
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 8, fig.height = 12, dev.args = list(type = "cairo"), fig.align = "center"}
ggdraw(p_done) + 
  draw_plot(hypoimg:::hypo_logo(.4), .1, .1, .8, .8) +
  draw_label("docs version", color = rgb(0,0,0,.2), size = 35, angle = 45)
```

Finally, we can export Figure S5.

```{r, eval = FALSE}
# export final figure
hypo_save(filename = 'figures/SF5.png',
          plot = p_done,
          width = 8,
          height = 12,
          type = "cairo",
          comment = plot_comment)
```

---
