---
output: html_document
editor_options:
  chunk_output_type: console
---

# Figure 2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')
library(GenomicOriginsScripts)
library(hypoimg)
library(patchwork)
args <- c('2_analysis/msmc/output/', '2_analysis/cross_coalescence/output/',
'2_analysis/msmc/setup/msmc_grouping.txt', '2_analysis/msmc/setup/msmc_cc_grouping.txt',
'2_analysis/summaries/fst_globals.txt')
```

## Summary

This is the accessory documentation of Figure 2.
The Figure can be recreated by running the **R** script `plot_F2.R`:
```sh
cd $BASE_DIR

Rscript --vanilla R/fig/plot_F2.R \
   2_analysis/msmc/output/ 2_analysis/cross_coalescence/output/ \
   2_analysis/msmc/setup/msmc_grouping.txt 2_analysis/msmc/setup/msmc_cc_grouping.txt \
   2_analysis/summaries/fst_globals.txt

```

## Details of `plot_F2.R`

In the following, the individual steps of the R script are documented.
It is an executable R script that depends on the accessory **R** packages [**GenomicOriginsScripts**](https://k-hench.github.io/GenomicOriginsScripts), as well as on the **R** packages [**hypoimg**](https://k-hench.github.io/hypoimg) and [**patchwork**](https://patchwork.data-imaginist.com/).

### Config

The scripts start with a header that contains copy & paste templates to execute or debug the script:

```{r, eval = FALSE}
#!/usr/bin/env Rscript
# run from terminal:
# Rscript --vanilla R/fig/plot_F2.R \
#    2_analysis/msmc/output/ 2_analysis/cross_coalescence/output/ \
#    2_analysis/msmc/setup/msmc_grouping.txt 2_analysis/msmc/setup/msmc_cc_grouping.txt \
#    2_analysis/summaries/fst_globals.txt
# ===============================================================
# This script produces Figure 2 of the study "Ancestral variation, hybridization and modularity
# fuel a marine radiation" by Hench, Helmkampf, McMillan and Puebla
# ---------------------------------------------------------------
# ===============================================================
# args <- c('2_analysis/msmc/output/', '2_analysis/cross_coalescence/output/',
# '2_analysis/msmc/setup/msmc_grouping.txt', '2_analysis/msmc/setup/msmc_cc_grouping.txt',
# '2_analysis/summaries/fst_globals.txt')
# script_name <- "R/fig/plot_F2.R"
# ----------------------------------------
```

The next section processes the input from the command line (supposedly...).
It stores the arguments in the vector `args`.
The **R** packages **GenomicOriginsScripts** and **patchwork** are loaded and the script name and the current working directory are stored inside variables (`script_name`, `plot_comment`).
This information will later be written into the meta data of the figure to help us tracing back the scripts that created the figures in the future.

Then we drop all the imported information besides the arguments following the script name and print the information to the terminal.

```{r, eval = FALSE}
args <- commandArgs(trailingOnly = FALSE)
# setup -----------------------
library(GenomicOriginsScripts)
library(hypoimg)
library(patchwork)

cat('\n')
script_name <- args[5] %>%
  str_remove(.,'--file=')

plot_comment <- script_name %>%
  str_c('mother-script = ',getwd(),'/',.)

args <- process_input(script_name, args)
```


```r
#> ── Script: scripts/plot_F2.R ────────────────────────────────────────────
#> Parameters read:
#> ★ 1: 2_analysis/msmc/output/
#> ★ 2: 2_analysis/cross_coalescence/output/
#> ★ 3: 2_analysis/msmc/setup/msmc_grouping.txt
#> ★ 4: 2_analysis/msmc/setup/msmc_cc_grouping.txt
#> ★ 5: 2_analysis/summaries/fst_globals.txt
#> ─────────────────────────────────────────── /current/working/directory ──
```

The directories for the demographic inference and the cross-coalescence data are received and stored in respective variables.
Also, the files containing the groupings for demographic inference and cross-coalescence as well as the reference file for the genome wide $F_{ST}$ values are received.

```{r, message = FALSE, warning = FALSE}
# config -----------------------
msmc_path <- as.character(args[1])
cc_path <- as.character(args[2])
msmc_group_file <- as.character(args[3])
cc_group_file <- as.character(args[4])
fst_globals_file <- as.character(args[5])
```

The **msmc** sample groupings are imported and the $F_{ST}$ values loaded.

```{r, message = FALSE, warning = FALSE}
# actual script =========================================================
msmc_groups <- read_tsv(msmc_group_file)
cc_groups <- read_tsv(cc_group_file)
fst_globals <- vroom::vroom(fst_globals_file,delim = '\t',
                            col_names = c('loc','run_prep','mean_fst','weighted_fst')) %>%
  separate(run_prep,into = c('pop1','pop2'),sep = '-') %>%
  mutate(run = str_c(pop1,loc,'-',pop2,loc),
         run = fct_reorder(run,weighted_fst))
```

Next, the file names of all **msmc** results are collected.

```{r, message = FALSE, warning = FALSE}
# locate cross-coalescence results
msmc_files <- dir(msmc_path, pattern = '.final.txt.gz')
cc_files <- dir(cc_path, pattern = '.final.txt.gz')
```

Separately, all the  demographic inference and cross-coalescence data are read in an compiled into two data sets.

```{r, message = FALSE, warning = FALSE}
# import effective population size data
msmc_data <- msmc_files %>%
  map_dfr(.f = get_msmc, msmc_path = msmc_path)

# import cross-coalescence data
cc_data <- cc_files %>%
  map_dfr(get_cc, cc_groups = cc_groups, cc_path = cc_path) %>%
  mutate( run = factor(run, levels = levels(fst_globals$run)))
```

The default color scheme is adjusted (to keep _H. unicolor_ visible) and the tick color for the plots is defined.

```{r, message = FALSE, warning = FALSE}
# color adjustments for line plots (replace white by gray)
clr_alt <- clr
clr_alt['uni'] <- rgb(.8,.8,.8)
clr_ticks <- 'lightgray'
```

The first panel containing the demographic history is created.

```{r, message = FALSE, warning = FALSE}
p_msmc <- msmc_data %>%
  # remove the two first and last time segments
  filter(!time_index %in% c(0:2,29:31)) %>%
  ggplot( aes(x=YBP, y=Ne, group = run_nr, colour = spec)) +
  # add guides for the logarithmic axes
  annotation_logticks(sides="tl", color = clr_ticks) +
  # add the msmc data as lines
  geom_line()+
  # set the color scheme
  scale_color_manual('Species',
                     values = clr_alt, label = sp_labs) +
  # format the x axis
  scale_x_log10(expand = c(0,0),
                breaks = c(10^3, 10^4, 10^5),
                position = 'top',
                labels = c("1-3 kya", "10-30 kya", "100-300 kya"),
                name = "Years Before Present") +
  # format the y axis
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)),
                breaks = c(10^3,10^4,10^5,10^6)) +
  # format the color legend
  guides(colour = guide_legend(title.position = "top",
                               override.aes = list(alpha = 1, size=1),
                               nrow = 2, byrow=TRUE)) +
  # set the axis titles
  labs(x="Generations Before Present",
       y = expression(Effective~Population~Size~(italic(N[e])))) +
  # set plot range
  coord_cartesian(xlim = c(250, 5*10^5)) +
  # tune plot appreance
  theme_minimal()+
  theme(axis.ticks = element_line(colour = clr_ticks),
        legend.position = c(1,.03),
        legend.justification = c(1,0),
        legend.text.align = 0,
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        title = element_text(face = 'bold'),
        axis.title = element_text(face = 'plain'),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        legend.title = element_text(face = 'plain'))
```

<center>
```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 11.1, fig.height = 5, dev.args = list(type = "cairo")}
p_msmc
```
</center>

The second panel containing the cross-coalescence is created.

```{r, message = FALSE, warning = FALSE}
p_cc <- cc_data %>%
  # remove the two first and last time segments
  filter( !time_index %in% c(0:2,29:31)) %>%
  arrange(run_nr) %>%
  # attach fst data
  left_join(fst_globals %>%
              select(run, weighted_fst)) %>%
  ggplot(aes(x = YBP, y = Cross_coal, group = run_nr, color = weighted_fst)) +
  # add guides for the logarithmic axis
  annotation_logticks(sides="b", color = clr_ticks) +
  # add the msmc data as lines
  geom_line(alpha = 0.2)+
  # set the color scheme
  scale_color_gradientn(name = expression(Global~weighted~italic(F[ST])),
                        colours = hypogen::hypo_clr_LGs[1:24])+
  # format the x axis
  scale_x_log10(expand = c(0,0),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  # set the axis titles
  guides(color = guide_colorbar(barheight = unit(7, 'pt'),
                                barwidth = unit(150, 'pt'),
                                title.position = 'top'))+
  # set the axis titles
  labs(x = "Generations Before Present",
       y = 'Cross-coalescence Rate') +
  # set plot range
  coord_cartesian(xlim = c(250, 5*10^5)) +
  # tune plot appreance
  theme_minimal()+
  theme(axis.ticks = element_line(colour = clr_ticks),
        legend.position = c(1,.03),
        legend.direction = 'horizontal',
        legend.justification = c(1,0),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        title = element_text(face = 'bold'),
        axis.title = element_text(face = 'plain'),
        legend.title = element_text(face = 'plain'))
```

Then, the figure is composed from both panels.

```{r, message = FALSE, warning = FALSE}
# combine panels a and b
p_done <- p_msmc / p_cc + plot_annotation(tag_levels = c('a'))
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 11.1, fig.height = 10.3, dev.args = list(type = "cairo"), fig.align = "center"}
p_done
```

Finally, we can export Figure 2.

```r
# export figure 2
hypo_save(plot = p_done, filename = 'figures/F2.pdf',
          width = 10, height = 7,
          comment = plot_comment,
          device = cairo_pdf)
```

---
