---
output: html_document
editor_options:
  chunk_output_type: console
---
# Supplementary Figure 5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      comment = "#>",
                      dev = "ragg_png",
                      eval = FALSE)
knitr::opts_knit$set(root.dir = '../')

library(ragg)
```


```{r , include = FALSE}
args <- c('2_analysis/fst/50k/', '2_analysis/summaries/fst_globals.txt')

library(GenomicOriginsScripts)
library(ggforce)
library(hypoimg)
library(hypogen)
library(vroom)

data_dir <- as.character(args[1])
globals_file <- as.character(args[2])
```

## Summary

This is the accessory documentation of Figure S5.
The Figure can be recreated by running the **R** script `plot_SF5.R`:

```sh
cd $BASE_DIR

Rscript --vanilla R/fig/plot_SF5.R \
    2_analysis/fst/50k/ \
    2_analysis/summaries/fst_globals.txt
```

## Details of `plot_SF5.R`

In the following, the individual steps of the R script are documented.
It is an executable R script that depends on the accessory R package [**GenomicOriginsScripts**](https://k-hench.github.io/GenomicOriginsScripts), as well as on the packages [**hypoimg**](https://k-hench.github.io/hypoimg), [**hypogen**](https://k-hench.github.io/hypogen) and [**patchwork**](https://patchwork.data-imaginist.com/)

### Config

The scripts start with a header that contains copy & paste templates to execute or debug the script:

```{r, eval = FALSE}
#!/usr/bin/env Rscript
# run from terminal:
# Rscript --vanilla R/fig/plot_SF5.R \
#     2_analysis/fst/50k/ \
#     2_analysis/summaries/fst_globals.txt
# ===============================================================
# This script produces Suppl. Figure 5 of the study "Rapid radiation in a
# highly diverse marine environment" by Hench, Helmkampf, McMillan and Puebla
# ---------------------------------------------------------------
# ===============================================================
# args <- c('2_analysis/fst/50k/', '2_analysis/summaries/fst_globals.txt')
# script_name <- "R/fig/plot_SF5.R"
args <- commandArgs(trailingOnly = FALSE)
```

The next section processes the input from the command line.
It stores the arguments in the vector `args`.
The needed R packages are loaded and the script name and the current working directory are stored inside variables (`script_name`, `plot_comment`).
This information will later be written into the meta data of the figure to help us tracing back the scripts that created the figures in the future.

Then we drop all the imported information besides the arguments following the script name and print the information to the terminal.

```{r, eval = FALSE}
# setup -----------------------
renv::activate()
library(GenomicOriginsScripts)
library(ggforce)
library(hypoimg)
library(hypogen)
library(vroom)

cat('\n')
script_name <- args[5] %>%
  str_remove(.,'--file=')

plot_comment <- script_name %>%
  str_c('mother-script = ',getwd(),'/',.)

args <- process_input(script_name, args)
```

```r
#> ── Script: R/fig/plot_SF5.R ────────────────────────────────────────────
#> Parameters read:
#> ★ 1: 2_analysis/fst/50k/
#> ★ 2: 2_analysis/summaries/fst_globals.txt
#> ────────────────────────────────────────── /current/working/directory ──
```

The directory containing the PCA data is received and stored in a variable.
Also the default color scheme is updated and the size of the hamlet ann.

```{r, message = FALSE, warning = FALSE}
# config -----------------------
data_dir <- as.character(args[1])
globals_file <- as.character(args[2])
```


```{r, message = FALSE, warning = FALSE}
# script -----------------------

# locate data files
files <- dir(path = data_dir, pattern = '.50k.windowed.weir.fst.gz')
```


```{r, message = FALSE, warning = FALSE}
# load genome wide average fst data
globals <- vroom::vroom(globals_file, delim = '\t',
                        col_names = c('loc','run','mean','weighted')) %>%
  mutate(run = str_c(loc,'-',run) %>%
           reformat_run_name()
  )
```


```{r, message = FALSE, warning = FALSE}
# prepare data import settings within a data table (tibble)
import_table <- list(file = str_c(data_dir,files),
                     fst_threshold = c(.5,.4,.3,.2,.1,.05,.02,.01)) %>%
  cross_df() %>%
  mutate( run =  file %>%
            str_remove('^.*/') %>%
            str_sub(.,1,11) %>%
            reformat_run_name())
```


```{r, message = FALSE, warning = FALSE}
# import dxy data and compute threshold stats
get_fst_fixed <- function(file, run, fst_threshold,...){

  data <- hypogen::hypo_import_windows(file, ...) %>%
    mutate(rank = rank(WEIGHTED_FST, ties.method = "random"))%>%
    mutate(thresh = fst_threshold) %>%
    mutate(outl = (WEIGHTED_FST > thresh) %>% as.numeric()) %>%
    filter(outl == 1 )

  if(nrow(data) == 0){
    return(tibble(run = run, n = 0, avg_length = NA, med_length = NA, min_length = NA, max_length = NA,
                  sd_length = NA, overal_length = NA, threshold_value = fst_threshold))
  } else {
    data %>%
      # next, we want to collapse overlapping windows
      group_by(CHROM) %>%
      # we check for overlap and create 'region' IDs
      mutate(check = 1-(lag(BIN_END,default = 0)>BIN_START),
                    ID = str_c(CHROM,'_',cumsum(check))) %>%
      ungroup() %>%
      # then we collapse the regions by ID
      group_by(ID) %>%
      summarise(run = run[1],
                       run = run[1],
                       treshold_value = thresh[1],
                       CHROM = CHROM[1],
                       BIN_START = min(BIN_START),
                       BIN_END = max(BIN_END)) %>%
      mutate(PEAK_SIZE = BIN_END-BIN_START) %>%
      summarize(run = run[1],
                       run = run[1],
                       n = length(ID),
                       avg_length = mean(PEAK_SIZE),
                       med_length = median(PEAK_SIZE),
                       min_length = min(PEAK_SIZE),
                       max_length = max(PEAK_SIZE),
                       sd_length = sd(PEAK_SIZE),
                       overal_length = sum(PEAK_SIZE),
                       threshold_value = treshold_value[1])
  }
}
```


```{r, message = FALSE, warning = FALSE}
# load data and compute statistics based on fixed fst threshold
data <- purrr::pmap_dfr(import_table, get_fst_fixed) %>%
  left_join(globals) %>%
  mutate(run = fct_reorder(run, weighted))
```


```{r, message = FALSE, warning = FALSE}
# pre-format labels
data2 <- data %>%
  select(threshold_value,weighted,n,avg_length,overal_length) %>%
  mutate(avg_length = avg_length/1000,
         overal_length = overal_length/(10^6)) %>%
  rename(`atop(Number~of,Regions)` = 'n',
         `atop(Average~Region,Length~(kb))` = 'avg_length',
         `atop(Cum.~Region,Length~(Mb))` = 'overal_length') %>%
  pivot_longer(names_to = 'variable',values_to = 'Value',3:5) %>%
  mutate(threshold_value = str_c('italic(F[ST])~threshold:~',
                                 threshold_value),
         variable = factor(variable, levels = c('atop(Number~of,Regions)',
                                                'atop(Average~Region,Length~(kb))',
                                                'atop(Cum.~Region,Length~(Mb))')))
```


```{r, message = FALSE, warning = FALSE}
# set font size
base_line_clr <- "black"
```


```{r, message = FALSE, warning = FALSE}
# compile plot
p_done <- data2 %>%
  # select thresholds of interest
  filter(!(threshold_value %in% (c(0.02,.1,0.2, 0.3, .4) %>%
                                   str_c("italic(F[ST])~threshold:~",.)))) %>%
  ggplot(aes(x = weighted, y = Value#, fill = weighted
             )
         )+
  # add red line for genome extent in lowest row
  geom_hline(data = tibble(variable = factor(c('atop(Cum.~Region,Length~(Mb))',
                                               'atop(Average~Region,Length~(kb))',
                                               'atop(Number~of,Regions)'),
                                             levels = c('atop(Number~of,Regions)',
                                                        'atop(Average~Region,Length~(kb))',
                                                        'atop(Cum.~Region,Length~(Mb))')),
                           y = c(559649677/(10^6),NA,NA)),
             aes(yintercept = y),
             color = rgb(1,0,0,.25))+
  # add data points
  geom_point(size = plot_size,
             color = plot_clr )+
  # define plot stucture
  facet_grid(variable~threshold_value,
             scale='free',
             switch = 'y',
             labeller = label_parsed)+
  # configure scales
  scale_x_continuous(name = expression(Whole-genome~differentiation~(weighted~italic(F[ST]))),
                     breaks = c(0,.05,.1),
                     limits = c(-.00025,.10025),
                     labels = c("0", "0.05", "0.1"))+
  # configure legend
  guides(fill = guide_colorbar(barwidth = unit(150, "pt"),
                               label.position = "top",
                               barheight = unit(5,"pt")))+
  # tweak plot apperance
  theme_minimal()+
  theme(axis.text = element_text(size = plot_text_size_small,
                                 color = rgb(.6,.6,.6)),
        axis.title.y = element_blank(),
        axis.text.x = element_text(vjust = .5, angle = 0),
        axis.title.x = element_text(vjust = -2),
        panel.background = element_rect(fill = rgb(.95,.95,.95,.5),
                                        color = rgb(.9,.9,.9,.5),
                                        size = .3),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(size = plot_lwd),
        legend.position = "bottom",
        strip.text = element_text(size = plot_text_size),
        legend.direction = "horizontal",
        strip.placement = 'outside',
        axis.title = element_text(size = plot_text_size),
        legend.title = element_text(size = plot_text_size),
        strip.background.y = element_blank(),
        plot.background = element_blank())
```

Finally, we can export Figure S5.

```{r, eval = FALSE}
# export figure 5
hypo_save(filename = 'figures/SF5.pdf',
          plot = p_done,
          width = f_width,
          height = .5 * f_width,
          device = cairo_pdf,
          comment = plot_comment,
          bg = "transparent")
```
